---
title: "Title of Project"
author: "510611635"
subtitle: "Project 3"
date: "University of Sydney | DATA1901 | May 2022"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

# Recommendation

# Evidence

```{r, message=F}
# Load packages
library(tidyverse)

# Read in the dataset (from the url or directly from the file in your folder)
toilets = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/projects/2022data/toilets.csv")
broadband = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/projects/2022data/broadband.csv", na.strings = c("", "NA"))
RBT = read.csv("RBT.csv")
```

```{r eval=FALSE, include=FALSE}
str(RBT)

# parse character into number
RBT$RBT.conducted = parse_number(RBT$RBT.conducted)
RBT$Positive.RBT = parse_number(RBT$Positive.RBT)
RBT$Licences = parse_number(RBT$Licences)
# parse int into date

# add column total death
RBT$totalDeath = RBT$Number.of.drivers.and.motorcycle.riders.killed.with.a.blood.alcohol.concentration..BAC..above.the.legal.limit + RBT$Number.of.deaths.from.crashes.involving.a.driver.or.motorcycle.rider.with.a.blood.alcohol.concentration..BAC..above.the.legal.limit
```

<!-- ## NSW: Positive RBT vs total death -->
```{r eval=FALSE, include=FALSE}
NSW = filter(RBT, State == "NSW")
# omit rows with na
NSW2 = NSW[1:12,]
# scatter plot: Positive RBT vs total death
ggplot(NSW2, aes(x = Positive.RBT, y = totalDeath)) + 
  geom_point() + geom_smooth(method = "lm")
# correlation coefficient
cor(NSW2$Positive.RBT, NSW2$totalDeath)
# linear model
fit = lm(NSW$totalDeath~NSW$Positive.RBT)
# residual plot
ggplot(NSW2, aes(x = Positive.RBT, y = fit$residuals)) + 
  geom_point() + geom_hline(yintercept = 0)
```

<!-- ## NSW: Positive RBT vs RBT conducted(low correlation) -->
```{r eval=FALSE, include=FALSE}
NSW = filter(RBT, State == "NSW")

# scatter plot: Positive RBT vs total death
ggplot(NSW, aes(x = RBT.conducted, y = Positive.RBT)) + 
  geom_point() + geom_smooth(method = "lm")
# correlation coefficient
cor(NSW$Positive.RBT, NSW$RBT.conducted)
# linear model
fit = lm(NSW$totalDeath~NSW$Positive.RBT)
# residual plot
ggplot(NSW2, aes(x = Positive.RBT, y = fit$residuals)) + 
  geom_point() + geom_hline(yintercept = 0)
```

## IDA
The dataset was published by Australian Competition and Consumer Commission at December 7th, 2021. The test was conducted during the Septemper 2021.
```{r IDA, }
# remove rows with missing value or impaired service
broadband <- broadband %>% na.omit() %>% filter(is_this_service_impaired == FALSE)
glimpse(broadband)
broadband %>% filter(is_this_service_underperforming == TRUE)
```

### The Retail Servive Providers in the Dataset 
```{r Retail Service Providers barplot}
# RSP and state
ggplot(broadband, aes(fct_infreq(rsp))) + geom_bar(aes(fill = state_or_territory)) +  scale_x_discrete(guide = guide_axis(n.dodge=2)) + labs(x = "tiers") + scale_fill_manual(values=c("#FF9A6B", "#F7FF66", "#fbcd69", "#47ff3d", "#00db66", "#00a85a", "#00754e"))
```


### tiers of services
```{r}
# tiers
broadband %>% filter(tier == "250/25 Mbps") %>% select(rsp) %>% table()
broadband %>% filter(tier == "25/5 Mbps") %>% select(rsp) %>% table()
# bar plot
ggplot(broadband, aes(factor(tier, level = c("Fixed Wireless Plus", "12/1 Mbps", "25/5 Mbps", "25/10 Mbps", "50/20 Mbps", "75/20 Mbps", "100/20 Mbps", "100/40 Mbps", "250/25 Mbps", "250/100 Mbps", "500-990/50 Mbps")))) + geom_bar(aes(fill = rsp)) + scale_x_discrete(guide = guide_axis(n.dodge=2)) + labs(x = "tiers")
table(broadband$tier)
```


## Reseatch Questions

### 50/20 Mbps tier analysis
```{r}
# Telstra vs others
tier_50_20 = broadband[broadband$tier == "50/20 Mbps", ]
telstra_tier_50_20 = tier_50_20[tier_50_20$rsp == "Telstra", ]
others_tier_50_20 = tier_50_20[tier_50_20$rsp != "Telstra", ]

```

#### All hour trimmed mean download speed
```{r}
# boxplot of all hour trimmed mean download speed by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.download.speed, fill = fct_reorder(rsp, All.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps All Hour Trimmed Mean Download Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean download
```{r}
# boxplot of busy hour trimmed mean download speed by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.download.speed, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps Busy Hour Trimmed Mean Download Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Paired T Test of trimmed mean download speed
```{r}
telstra_tier_50_20_download_diff = telstra_tier_50_20$All.hour.trimmed.mean.download.speed - telstra_tier_50_20$Busy.hour.trimmed.mean.download.speed
t.test(telstra_tier_50_20_download_diff, mu = 0)
```


#### All hour trimmed mean upload
```{r}
# boxplot of all hour trimmed mean Upload speed by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.upload.speed, fill = fct_reorder(rsp, All.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps All Hour Trimmed Mean Upload Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean upload
```{r}
# boxplot of busy hour trimmed mean upload speed by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.upload.speed, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps Busy Hour Trimmed Mean Upload Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### All hour trimmed mean latency
```{r}
# boxplot of all hour trimmed mean Latency by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.latency, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.latency, fill = fct_reorder(rsp, All.hour.trimmed.mean.latency, .fun = median, .desc = FALSE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps All Hour Trimmed Mean Latency by RSP", x = "Retail Service Providers", y = "ms") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean latency
```{r}
# boxplot of busy hour trimmed mean Latency by RSP
tier_50_20 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.latency, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.latency, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.latency, .fun = median, .desc = FALSE))) + geom_boxplot() + 
  labs(title = "Tier 50/20 Mbps Busy Hour Trimmed Mean Latency by RSP", x = "Retail Service Providers", y = "ms") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Speed comparison
```{r}
speed = data.frame(
  Categories = c("All hour download speed", "All hour upload speed", "All hour download speed", "All hour upload speed", 
                 "Busy hour download speed", "Busy hour upload speed", "Busy hour download speed", "Busy hour upload speed"),
  Mbps = c(telstra_tier_50_20$All.hour.trimmed.mean.download.speed %>% median(), telstra_tier_50_20$All.hour.trimmed.mean.upload.speed %>% median(), others_tier_50_20$All.hour.trimmed.mean.download.speed %>% median(), others_tier_50_20$All.hour.trimmed.mean.upload.speed %>% median(), 
           telstra_tier_50_20$Busy.hour.trimmed.mean.download.speed %>% median(), telstra_tier_50_20$Busy.hour.trimmed.mean.upload.speed %>% median(), others_tier_50_20$Busy.hour.trimmed.mean.download.speed %>% median(), others_tier_50_20$Busy.hour.trimmed.mean.upload.speed %>% median()),
  RSP = c("Telstra", "Telstra", "Others", "Others", "Telstra", "Telstra", "Others", "Others")
)
ggplot(speed, aes(Categories, Mbps, fill = RSP)) + 
  geom_col(position = "dodge" ) # + scale_fill_manual(values=c("#FF9A6B", "#F7FF66"))
```

#### Latency comparison
```{r}
latency = data.frame(
  Categories = c("All hour latency", "Busy hour latency", "All hour latency", "Busy hour latency"),
  ms = c(telstra_tier_50_20$All.hour.trimmed.mean.latency %>% median(), others_tier_50_20$All.hour.trimmed.mean.latency %>% median(), 
           telstra_tier_50_20$Busy.hour.trimmed.mean.latency %>% median(), others_tier_50_20$Busy.hour.trimmed.mean.latency %>% median()),
  RSP = c("Telstra", "Telstra", "Others", "Others")
)
ggplot(latency, aes(Categories, ms, fill = RSP)) + 
  geom_col(position = "dodge" ) # + scale_fill_manual(values=c("#FF9A6B", "#F7FF66"))

```

---

### 100/40 Mbps tier analysis
```{r}
# Telstra vs others
tier_100_40 = broadband[broadband$tier == "100/40 Mbps", ]
telstra_tier_100_40 = tier_100_40[tier_100_40$rsp == "Telstra", ]
others_tier_100_40 = tier_100_40[tier_100_40$rsp != "Telstra", ]
```

#### All hour trimmed mean download
```{r}
# boxplot of all hour trimmed mean download speed by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.download.speed, fill = fct_reorder(rsp, All.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 100/40 Mbps All Hour Trimmed Mean Download Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean download
```{r}
# boxplot of busy hour trimmed mean download speed by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.download.speed, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.download.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 100/40 Mbps Busy Hour Trimmed Mean Download Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### All hour trimmed mean upload
```{r}
# boxplot of all hour trimmed mean upload speed by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.upload.speed, fill = fct_reorder(rsp, All.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 100/40 Mbps All Hour Trimmed Mean Upload Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean upload
```{r}
# boxplot of busy hour trimmed mean upload speed by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.upload.speed, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.upload.speed, .fun = median, .desc = TRUE))) + geom_boxplot() + 
  labs(title = "Tier 100/40 Mbps Busy Hour Trimmed Mean Upload Speed by RSP", x = "Retail Service Providers", y = "Mbps") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### All hour trimmed mean latency
```{r}
# boxplot of all hour trimmed mean Latency by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, All.hour.trimmed.mean.latency, .fun = median, .desc = TRUE), y = All.hour.trimmed.mean.latency, fill = fct_reorder(rsp, All.hour.trimmed.mean.latency, .fun = median, .desc = FALSE))) + geom_boxplot() + 
  labs(title = "Tier 100/40 Mbps All Hour Trimmed Mean Latency by RSP", x = "Retail Service Providers", y = "ms") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

#### Busy hour trimmed mean latency
```{r}
# boxplot of busy hour trimmed mean Latency by RSP
tier_100_40 %>% ggplot(aes(x = fct_reorder(rsp, Busy.hour.trimmed.mean.latency, .fun = median, .desc = TRUE), y = Busy.hour.trimmed.mean.latency, fill = fct_reorder(rsp, Busy.hour.trimmed.mean.latency, .fun = median, .desc = F))) + geom_violin() + 
  labs(title = "Tier 100/40 Mbps Busy Hour Trimmed Mean Latency by RSP", x = "Retail Service Providers", y = "ms") + scale_fill_discrete(guide = guide_legend(title = "RSP")) + scale_x_discrete(guide = guide_axis(n.dodge=2)) + coord_flip()
```

#  Acknowledgments